---
title: ""
format:
  html:
    toc: false
---

![](headshot.svg){fig-align="center" width="225"}

::: {.text .padding .fade-in}
Hi 👋 I'm a Research Fellow at the Gambling Treatment & Research Centre, University of Sydney, Australia.

You can find out more about me [here](https://robheirene.netlify.app/about.html).
:::

------------------------------------------------------------------------

<br>

::: fade-in
## Latest research findings

[The state of the open science in the behavioural addictions field]{style="color: #289998; font-weight: bold;"}

I recently contributed to a conceptual paper discussing the landscape of open science in behavioural addictions research---find out more [here](https://robheirene.netlify.app/publication/2023-open-science-landscape-eben-heirene/). The paper was led by [Dr Charlotte Eben](https://twitter.com/CharlotteEben), who is doing some great work in this space!

## What I'm working on

[Moving accross the world (again)]{style="color: #289998; font-weight: bold;"}

As of November 2023, Im leaving the University of Plymouth in the UK and heading back to the University of Sydney again to finally settle in a place that feels like home. Doing that with a family of six (four young children) is keeping me pretty busy right now...

[Pondering + plotting]{style="color: #289998; font-weight: bold;"}

I've also thinking about how to progress gambling harm prevention in my research and enjoying learning how to make interactive charts using the `plotly` R package. I'm marrying these interests by making the below chart of prevention strategies for problem gambling.

[**Click on the bars to see more**]{style="color: #289998; font-weight: bold;"}. It's best viewed on a pc/mac or tablet (if you are on mobile, turn it to landscape mode).
:::

```{r echo =FALSE, results=FALSE, warning=FALSE, message=FALSE}
# I keep this in just in case I want to run any code locally
# Install and load the groundhog package to ensure consistency of the package versions used here:
# install.packages("groundhog") # Install

# library(groundhog) # Load

# packages <- c('tidyverse', # Clean, organise, and visualise data
              # 'plotly', # Add interactive elements to figures
              # 'sysfonts', # Special fonts for figures
              # 'showtext', # Special fonts for figures
              # 'htmlwidgets') # Make plotly plots HTML format for rendering in Quarto
              
# Load desired package with versions specific to project start date:
# groundhog.library(packages, "2023-06-05") # Set Slightly earlier than start date as some of the plotly dependency packages wouldn't load properly with date as 07/07/2023

# Load new font for figures/graphs
# font_add_google("Poppins")
# showtext_auto()
```

```{r echo =FALSE,warning=FALSE, message=FALSE}

```

```{=html}
<iframe src="posts/2023-gambling-harms-chart/sunburst_fig.html" class="fade-inhtml" width="100%" height="750" style="border:none;"></iframe>
```
::: fade-in
It's very much a work in progress (*I still need to add many strategies and labels for those I have*), but you can find out more about my plans for this [here](https://robheirene.netlify.app/posts/2023-gambling-harms-chart).
:::
