---
title: ""
format:
  html:
    toc: false
---

![](headshot.svg){fig-align="center" width="225"}

::: {.text .padding .fade-in}
Hi 👋 I'm a Lecturer in Psychology and Addiction Researcher at the [University of Plymouth](https://www.plymouth.ac.uk/staff/rob-heirene), UK.

You can find out more about me [here](https://robheirene.netlify.app/about.html).
:::

------------------------------------------------------------------------

<br>

::: fade-in
## Latest research findings

[The state of the open science in the behavioural addictions field]{style="color: #289998; font-weight: bold;"}

I recently contributed to a conceptual paper discussing the landscape of open science in behavioural addictions research---find out more [here](https://robheirene.netlify.app/publication/2023-open-science-landscape-eben-heirene/). The paper was led by [Dr Charlotte Eben](https://twitter.com/CharlotteEben), who is doing some great work in this space!

## What I'm working on

[Pondering + plotting]{style="color: #289998; font-weight: bold;"}

I'm currently thinking about how to progress gambling harm prevention in my research and enjoying learning how to make interactive charts using the `plotly` R package. I'm marrying these interests by making the below chart of prevention strategies for problem gambling.

[Click on the bars to see more]{style="color: #289998; font-weight: bold;"}. It's best viewed on a pc/mac or tablet, not on mobile (if you are on mobile, turn it to landscape mode).
:::

```{r echo =FALSE, results=FALSE, warning=FALSE, message=FALSE}
# Install and load the groundhog package to ensure consistency of the package versions used here:
# install.packages("groundhog") # Install

library(groundhog) # Load

packages <- c('tidyverse', # Clean, organise, and visualise data
              'plotly', # Add interactive elements to figures
              'sysfonts', # Special fonts for figures
              'showtext', # Special fonts for figures
              'htmlwidgets') # Make plotly plots HTML format for rendering in Quarto
              
# Load desired package with versions specific to project start date:
groundhog.library(packages, "2023-06-05") # Set Slightly earlier than start date as some of the plotly dependency packages wouldn't load properly with date as 07/07/2023

# Load new font for figures/graphs
font_add_google("Poppins")
showtext_auto()
```

```{r echo =FALSE,warning=FALSE, message=FALSE}
# Load manually input data:
# gam_prevention_data<- read.csv("posts/2023-gambling-harms-chart/gamb_prevention_data.csv") %>%
#   select(Layer:Description) # Warning, something as simple as a' and this dataset can cause it to not play nice
# 
# # Make sunburst chart:
# gam_prevention_sunburst<- plot_ly(gam_prevention_data,
#                                    ids = ~IDs,
#                                    labels = ~Label,
#                                    parents = ~Parents,
#                                    hoverinfo = "text",  # Needed for below argument to work
#                                    hovertext = ~Description,
#                                    type = 'sunburst',
#                                    maxdepth = 3,
#                                    insidetextorientation = 'radial',
#                                    marker = list(line = list(width = 5, color = '#002B36')),  # Adjust border size + color
#                                    sort = TRUE) %>%
#    plotly::layout(colorway = c("#289998", "#288D98", "#287699", "#7E6F98", "#BFB3C0", "#FFF7E7", "#E6C97F", "#8C8033", "#7E9F6E", "#65BBA3", "#94C8C0", "#998F28", "#62A899", "#986289", "#289970"),
#          font = list(
#            size = 14,
#            family = "Reem Kufi", # I can't get this to work so let me know if you can!
#                      color = "#FFF7E7"), # Set main font so the central layer text can be seen
#          paper_bgcolor = '#002B36', # background consistent with site
#          plot_bgcolor = '#002B36',
#          hoverlabel = list(font = list(size = 12, family = "Reem Kufi")))  # Increase the font size for hover text)
#  
# # The standard way of changing the font in ggplotly doesn't seem to work for me, and it seems like other people having the same issue. I found this workaround online (https://github.com/plotly/plotly.R/issues/2117) which I now use below to load and use the correct font in our ggplotly:
# 
# # Get the URL for the "Reem Kufi" font from Google Fonts:
# reem_kufi_file <- showtextdb::google_fonts("Reem Kufi")$regular_url 
# 
# # Create custom CSS:
# reem_kufi_css <- paste0(
#   "<style type = 'text/css'>",
#     "@font-face { ",
#       "font-family: 'Reem Kufi'; ", 
#       "src: url('", reem_kufi_file, "'); ",
#     "}",
#   "</style>"
# )
# 
# 
# # Add the CSS as a dependency for the plotly plot:
# gam_prevention_sunburst$dependencies <- c(
#   gam_prevention_sunburst$dependencies,
#   list(
#     htmltools::htmlDependency(
#       name = "reem-kufi-font", 
#       version = "0",  
#       src = "",
#       head = reem_kufi_css)))
# 
# # Save plot as a HTML file (it doesn't render well with quarto otherwise):
# saveWidget(gam_prevention_sunburst, 'fig.html')
```

```{=html}
<iframe src="posts/2023-gambling-harms-chart/fig.html" class="fade-inhtml" width="100%" height="750" style="border:none;"></iframe>
```
::: fade-in
It's very much a work in progress (*I still need to add many strategies and labels for those I have*), but you can find out more about my plans for this [here](https://robheirene.netlify.app/posts/2023-gambling-harms-chart).
:::
